Theories of deliberated choice

### 2022-10-26 simple.tex
Complicated remarks about shortcoming of this simple approach

An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
It is said to be truthful iff this claim is correct.
More precisely, its claims are about paths $\hist$ containing ${\arga, \ar_2, \ar_3, …, \ar_k}$ where $\ar_2 \in \dgi(\arga)$, $\ar_3 \in \dgi(\ar_2)$, …, $\ar_k \in \histend$. The following definition makes this precise.
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff $\forall i \in I, k \in \N^*, (\ar_j)_{j \in \intvl{1, k}} \in \allargs^k, \hist \in \allhist \suchthat [\ar_1 = \arga \land \ar_k \in \histend \land \forall j \in \intvl{1, k}: \ar_j \in R(\hist) \land \forall j \in \intvl{2, k}: \ar_j \in \dgi(\ar_{j - 1}) \land \hist \nileadsto \pga]: [d(\ar_k, \hist) ≠ \emptyset]$.
\end{definition}

>>>>

An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
It is said to be truthful iff this claim is correct.
More precisely, its claims are about paths $\hist$ containing ${\arga, \ar_2, \ar_3, …, \ar_k}$ where $\ar_2 \in \dgi(\arga)$, $\ar_3 \in \dgi(\ar_2)$, …, $\ar_k \in \histend$. The following definition makes this precise.
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff $\forall i \in I, k \in \N^*, (\ar_j)_{j \in \intvl{1, k}} \in \allargs^k, \hist \in \allhist \suchthat [\forall j \in \intvl{1, k}: \ar_j \in R(\hist) \land \ar_k \in \histend \land \forall j \in \intvl{2, k}: \ar_j \in \dgi(\ar_{j - 1}) \land \hist \nileadsto \pga]: [d(\ar_k, \hist) ≠ \emptyset]$.
\end{definition}

\begin{remark}
	There is a problem. First, truthfulness must be reduced as we can empirically only test $(a_1, a_5, a_2)$ when $a_2$ defends $a_1$ \emph{because of $a_5$} (not because of something else).
	Second, consider the following may hold for some $i$. $a_1, a_5, a_2$ not convincing, $a_1, a_4$ not convincing, whil $a_1, a_5$ and $a_1, a_4, a_2$ and other variants convincing. Then for all $i$ one of $a_1$ or $a_2$ will appear to be decisive although for those people neither is.

	Problem might be solved by observing that $(a_1, a_5, a_2, a_1)$ convincing means that $a_1$ is decisive after all; and not convincing means that we have observed not truthfulness.
\end{remark}
\begin{remark}
	We want that if the theory is invalid (thus some $i$ is not convinced), then either axiom fails or the theory is experimentally falsifiable.
	In the following example, the theory seems invalid but the axiom holds and the theory can’t seem to be defeated experimentally.

	Imagine $i$ thinks as follows. $a_1$ counter $a_5$; $a_2$ counters $a_4$; but $a_4$ and $a_5$ together make $a_1$ and $a_2$ invalid, so that $(a_4, a_5)$ counter both $a_1$ and $a_2$ (and also $a_1$ and $a_2$ together). Also assume that $i$ has short memory so that $a_1, a_4, a_5$ does not convince ($a_4$ and $a_5$ win) while $a_4, a_5, a_1$ do ($a_1$ wins as $a_4$ is too far away). BUT $a_5, a_4, a_1$ should then not convince.

	Path is convincing depending only the last two arguments (unordered): $\set{(a_1, a_5), (a_2, a_4), (a_1, a_2)}$ are convincing; $\set{(a_1, a_4), (a_2, a_5), (a_4, a_5)}$ are not.
	\begin{description}
		\item[$(a_1, a_2, a_4, a_5)$] $\nileadsto p$
		\item[$(a_1, a_2, a_5, a_4)$] $\nileadsto p$
		\item[$(a_1, a_4, a_2, a_5)$] $\nileadsto p$
		\item[$(a_1, a_4, a_5, a_2)$] $\nileadsto p$
		\item[$(a_1, a_5, a_2, a_4)$] $\ileadsto p$
		\item[$(a_1, a_5, a_4, a_2)$] $\ileadsto p$
		\item[$(a_2, a_1, a_4, a_5)$] $\nileadsto p$
		\item[$(a_2, a_1, a_5, a_4)$] $\nileadsto p$
		\item[$(a_2, a_4, a_1, a_5)$] $\ileadsto p$
		\item[$(a_2, a_4, a_5, a_1)$] $\ileadsto p$
		\item[$(a_2, a_5, a_1, a_4)$] $\nileadsto p$
		\item[$(a_2, a_5, a_4, a_1)$] $\nileadsto p$
		\item[$(a_4, a_1, a_2, a_5)$] $\nileadsto p$
		\item[$(a_4, a_1, a_5, a_2)$] $\nileadsto p$
		\item[$(a_4, a_2, a_1, a_5)$] $\ileadsto p$
		\item[$(a_4, a_2, a_5, a_1)$] $\ileadsto p$
		\item[$(a_4, a_5, a_1, a_2)$] $\ileadsto p$
		\item[$(a_4, a_5, a_2, a_1)$] $\ileadsto p$
		\item[$(a_5, a_1, a_2, a_4)$] $\ileadsto p$
		\item[$(a_5, a_1, a_4, a_2)$] $\ileadsto p$
		\item[$(a_5, a_2, a_1, a_4)$] $\nileadsto p$
		\item[$(a_5, a_2, a_4, a_1)$] $\nileadsto p$
		\item[$(a_5, a_4, a_1, a_2)$] $\ileadsto p$
		\item[$(a_5, a_4, a_2, a_1)$] $\ileadsto p$
	\end{description}
	It follows that no argument is decisive.
	The axiom' fails as $(a_4, a_1, a_5, a_2) \nileadsto p$ but $(a_4, a_1, a_5, a_4, a_2) \ileadsto p$!

	Let’s treat $a_1$ as including $a_5$ and $a_2$ as including $a_4$.

	Ends with $a_1$ and $a_2$: convincing. Ends with only $a_1$: convincing. Ends with $a_1$ and $a_4$: no. Ends with $a_2$ and $a_5$: no.
	\begin{description}
		\item[$(a_1, a_2, a_4, a_5)$] $\ileadsto p$
		\item[$(a_1, a_2, a_5, a_4)$] $\ileadsto p$
		\item[$(a_1, a_4, a_2, a_5)$] $\nileadsto p$
		\item[$(a_1, a_4, a_5, a_2)$] $\nileadsto p$
		\item[$(a_1, a_5, a_2, a_4)$] $\nileadsto p$
		\item[$(a_1, a_5, a_4, a_2)$] $\nileadsto p$
		\item[$(a_2, a_1, a_4, a_5)$] $\nileadsto p$
		\item[$(a_2, a_1, a_5, a_4)$] $\nileadsto p$
		\item[$(a_2, a_4, a_1, a_5)$] $\ileadsto p$
		\item[$(a_2, a_4, a_5, a_1)$] $\ileadsto p$
		\item[$(a_2, a_5, a_1, a_4)$] $\nileadsto p$
		\item[$(a_2, a_5, a_4, a_1)$] $\nileadsto p$
		\item[$(a_4, a_1, a_2, a_5)$] $\nileadsto p$
		\item[$(a_4, a_1, a_5, a_2)$] $\nileadsto p$
		\item[$(a_4, a_2, a_1, a_5)$] $\ileadsto p$
		\item[$(a_4, a_2, a_5, a_1)$] $\ileadsto p$
		\item[$(a_4, a_5, a_1, a_2)$] $\ileadsto p$
		\item[$(a_4, a_5, a_2, a_1)$] $\ileadsto p$
		\item[$(a_5, a_1, a_2, a_4)$] $\ileadsto p$
		\item[$(a_5, a_1, a_4, a_2)$] $\ileadsto p$
		\item[$(a_5, a_2, a_1, a_4)$] $\nileadsto p$
		\item[$(a_5, a_2, a_4, a_1)$] $\nileadsto p$
		\item[$(a_5, a_4, a_1, a_2)$] $\ileadsto p$
		\item[$(a_5, a_4, a_2, a_1)$] $\ileadsto p$
	\end{description}
	It follows that no argument is decisive.
\end{remark}
